---
title: "Install the model and all software/data required"
subtitle: MAgPIE model development team (magpie@pik-potsdam.de)
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Anastasis Giannousakis (giannou@pik-potsdam.de)
output:
  pdf_document: default
  toc: yes
  toc_float: true
  html_notebook: default
---

## HOW TO INSTALL
MAgPIE requires *GAMS* (https://www.gams.com/) including licenses for the
solvers *CONOPT* and (optionally) *CPLEX* for its core calculations. For the day
of the workshop all licenses will be provided as time-limited course licenses. 
As the model benefits significantly of recent improvements in *GAMS* and *CONOPT4* 
it is recommended to work with the most recent versions of both. 
Please make sure that the GAMS installation path is added to the PATH variable
of the system:

* the easiest way to add is by simply checking the "Use advanced installation mode" box at the beginning of the installation. At a later step you have to tick again a checkbox that adds the GAMS path to your PATH variable
* you can also edit your computer's advanced settings and add the GAMS path to the PATH variable manually (applies also if GAMS is installed but not included in PATH)

In addition *R* (https://www.r-project.org/) is required for pre- and
postprocessing and run management (needs to be added to the user's PATH variable
as well). It is recommended to install also RSudio (https://www.rstudio.com).

For R, some packages are required to run MAgPIE. All except of one (`gdxrrw`) are
either distributed via the offical R CRAN or via a separate repository hosted at
PIK (PIK-CRAN). Before proceeding PIK-CRAN should be added to the list of
available repositories via:
```{r, eval=FALSE}
options(repos = c(CRAN = "@CRAN@", pik = "https://rse.pik-potsdam.de/r/packages"))
```

The `gdxrrw` package has to be downloaded directly from GAMS via
```{r, eval=FALSE}
download.file("https://support.gams.com/_media/gdxrrw:gdxrrw_1.0.2.zip",
              "gdxrrw_1.0.2.zip")
```
In some cases it can happen that `gdxrrw` does not return an error message during
installation but also did not install properly. To verify a successful
installation try to load the package via `library(gdxrrw)`.

--------------------------------------------------------------------------------

If loading of the package fails you need to install the package from source.
Under Windows this requires to install Rtools
(https://cran.r-project.org/bin/windows/Rtools/) and to add it to the PATH
variable. After that you can run the following lines of code:

```{r, eval=FALSE}
download.file("https://support.gams.com/_media/gdxrrw:gdxrrw_1.0.2.tar.gz",
              "gdxrrw_1.0.2.tar.gz")
install.packages("gdxrrw_1.0.2.tar.gz",repos = NULL, type="source")
```

--------------------------------------------------------------------------------


After that all remaining packages can be installed via `install.packages`

```{r, eval=FALSE}
pkgs <- c("ggplot2",
          "curl",
          "gdx",
          "magclass",
          "madrat",
          "mip",
          "lucode",
          "magpie4",
          "magpiesets",
          "lusweave",
          "luscale",
          "goxygen")
install.packages(pkgs)
```

For post-processing model outputs *Latex* is required
(https://www.latex-project.org/get/). To be seen by the model it also needs to
be added to the PATH variable of your system.

If the following lines of code are executed withour error, then you are all set!
```{r, eval=FALSE}
system("gams")
library(gdxrrw)
library(magpie)
print("")
if(.Platform$OS.type == "unix") {
  system("which pdflatex")
} else {
  system("where pdflatex")
}
```
NOTE: If the model fails to start from the Windows console, try starting it from within RStudio.
