---
title: "Install the MAgPIE model and all software/data required"
subtitle: MAgPIE model development team (magpie@pik-potsdam.de)
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Anastasis Giannousakis (giannou@pik-potsdam.de)
output:
  html_notebook: default
  toc: yes
  toc_float: true
  pdf_document: default
---

## HOW TO INSTALL
There are two ways to download the MAgPIE model. If you have git installed clone the model from https://github.com/magpiemodel/magpie.git If you have no git installed you can alternatively download MAgPIE as a zip file from https://github.com/magpiemodel/magpie/archive/master.zip.

MAgPIE requires *GAMS* (https://www.gams.com/) including licenses for the
solvers *CONOPT* and (optionally) *CPLEX* for its core calculations. As the model benefits significantly from recent improvements in *GAMS* and *CONOPT4* 
it is recommended to work with the most recent versions of both. Please make sure that the GAMS installation path is added to the PATH variable
of the system:

* the easiest way to add is by simply checking the "Use advanced installation mode" box at the beginning of the installation. At a later step you have to tick again a checkbox that adds the GAMS path to your PATH variable
* you can also edit your computer's advanced settings and add the GAMS path to the PATH variable manually (applies also if GAMS is installed but not included in PATH).

This tutorial shows how to check and add variables to your PATH variable: https://www.youtube.com/watch?v=5P9EDJwfXBo

Please add the GAMS training license you have been provided (gamslice.txt) by saving the file to your GAMS local folder. Under Windows something like `C:\Program Files (x86)\GAMS\24.3`

In addition *R* (https://www.r-project.org/) is required for pre- and
postprocessing and run management (needs to be added to the user's PATH variable
as well). It is recommended to install also RSudio (https://www.rstudio.com).

For R, some packages are required to run MAgPIE. All except of one (`gdxrrw`) are
either distributed via the offical R CRAN or via a separate repository hosted at
PIK (PIK-CRAN). Before proceeding PIK-CRAN should be added to the list of
available repositories via:
```{r, eval=FALSE}
options(repos = c(CRAN = "@CRAN@", pik = "https://rse.pik-potsdam.de/r/packages"))
```

The `gdxrrw` package has to be downloaded directly from GAMS via
```{r, eval=FALSE}
download.file("https://support.gams.com/_media/gdxrrw:gdxrrw_1.0.2.zip",
              "gdxrrw_1.0.2.zip")
install.packages("gdxrrw_1.0.2.zip",repos = NULL)
```

It can be the case that the `R` installation reqiures a more up-to-date version of
`gdxrrw` package. In that case, check on the webpage https://support.gams.com/gdxrrw:interfacing_gams_and_r
for adequate package version and eventual additional information and then proceed 
with download and install.

In some cases it can happen that `gdxrrw` does not return an error message during
installation but also did not install properly. To verify a successful
installation try to load the package via `library(gdxrrw)`.

--------------------------------------------------------------------------------

If loading of the package fails you need to install the package from source.
Under Windows this requires to install Rtools
(https://cran.r-project.org/bin/windows/Rtools/) and to add it to the PATH
variable. After that you can run the following lines of code:

```{r, eval=FALSE}
download.file("https://support.gams.com/_media/gdxrrw:gdxrrw_1.0.2.tar.gz",
              "gdxrrw_1.0.2.tar.gz")
install.packages("gdxrrw_1.0.2.tar.gz",repos = NULL, type="source")
```

--------------------------------------------------------------------------------


After that all remaining packages can be installed via `install.packages`

```{r, eval=FALSE}
pkgs <- c("ggplot2",
          "curl",
          "gdx",
          "magclass",
          "madrat",
          "mip",
          "lucode",
          "magpie4",
          "magpiesets",
          "lusweave",
          "luscale",
          "goxygen",
          "shinyresults")
install.packages(pkgs)
```

For post-processing model outputs *Latex* is required
(https://www.latex-project.org/get/). To be seen by the model it also needs to
be added to the PATH variable of your system.

Lastly, please download MAgPIE from here: https://github.com/magpiemodel/magpie/archive/master.zip

If the following lines of code are executed withour error, then you are all set!
```{r, eval=FALSE}
system("gams")
library(gdxrrw)
library(magpie4)
print("")
if(.Platform$OS.type == "unix") {
  system("which pdflatex")
} else {
  system("where pdflatex")
}
```
NOTE: If the model fails to start from the Windows console, try starting it from within RStudio.
