---
title: "Advanced: Change MAgPIE GAMS Code"
subtitle: MAgPIE model development team (magpie@pik-potsdam.de)
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Florian Humpenöder (humpenoeder@pik-potsdam.de)
output:
  rmarkdown::github_document:
    toc: yes
---

# Learning objectives

MAgPIE has a modular concept. Each module (e.g. cropland) can have several realization (e.g. dynamic and static).
This tutorial shows how to add a new realization to a module in 4 steps:

1. create new realization via copying existing one
2. include it properly via lucode::update_modules_embedding()
3. do simple modification in new realization
4. choose realization in cfg and run model


# Getting started

We want to add a new realization to the cropland module.
In the MAgPIE 4.0 release the cropland module (30_crop) has only one realization called "endo_jun13". In this realization irrigation of bioenergy crops is prohibited. We will add a new realization, based on "endo_jun13", which allows for irrigation of bioenergy crops. We will call the new new realization "bioen_irrig".

## Step 1
We first copy-paste the folder "endo_jun13" and the file "endo_jun13.gms" (both located in "modules/30_crop"), and rename them to "bioen_irrig" and "bioen_irrig.gms" respectivly.

## Step 2
To include the new realization "bioen_irrig" properly into the GAMS code we run a specific R command in the main folder. First navigate in your command line to the MAgPIE main folder, then open a new R session (type "R" followed by ENTER), and then copy-paste the following R command:
```{r comment=NA,eval=FALSE}
lucode::update_modules_embedding()
```

## Step 3
The prohibition of irrigated bioenergy is specified in the file "presolve.gms" within the "bioen_irrig" realization.
We just have to remove (or comment out) lines 7-9 from this file to allow for irrigation of bioenergy crops. 

```{r comment=NA,eval=FALSE}
*vm_area.fx(j,"begr","irrigated")=0;
*vm_area.fx(j,"begr","irrigated")=0;
*vm_area.fx(j,"betr","irrigated")=0;
```

## Step 4
To test the new realization we have to make a change in the file "config/default.cfg" in line 218.
We change 
```{r comment=NA,eval=FALSE}
cfg$gms$crop    <- "endo_jun13"
```
to 
```{r comment=NA,eval=FALSE}
cfg$gms$crop    <- "bioen_irrig"
```
We can then start the model run with 
```{r comment=NA,eval=FALSE}
Rscript start.R -> 1: default -> 1: Direct execution
```